---
title: "2_CategoricalAnalysis"
output: html_document
---

## Purpose

Meaningfully reduce the number of categories for Water Column Position and Primary Habitat.
phylopath only works with binary categorical variables.

Also explore the correlation between Water Column Position and Habitat Use.

```{r}
# Load tidyverse for sanity
library(tidyverse)

# Load scico for plotting
library(scico)

# Load ape, geiger, nlme, and phytools for PGLS
library(ape)
library(geiger)
library(nlme)
library(phytools)
```

### Step 2: Load data

```{r}
# Load shark data
categoricalData <- read.csv("Data/SharkData/TraitData.csv")
```

### Step 3: Compare pelagic, benthopelagic, and benthic water column position

The first approach is to compare CFAR values for the three water column position levels:
* Benthic
* Benthopelagic
* Pelagic

We'll first visualize the spread of CFAR by water column position.

```{r}
FigureS2 <- ggplot() +
  geom_jitter(data = categoricalData, 
              aes(x = WaterColumnPosition, y = CFAR,
                  color = WaterColumnPosition), size = 2) +
  labs(x = element_blank(),
       y = "Caudal Fin Aspect Ratio (CFAR)") +
  scale_colour_manual(values = c("Pelagic" = "#82e1da", 
                               "Benthopelagic" = "#3b898e", 
                               "Benthic" = "#103f4d")) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.y = element_text(size = 12),
        legend.position = "none")

FigureS2

# Save plot
ggsave("Supplementary_Material/Results/FigS2.tiff",
       plot = FigureS2, units = "px",
       device = "tiff",
       width = 1200, height = 1300, dpi = 300)
```

Now we'll statistically evaluate which category benthopelagic should be binned with.

Biologically, we expect support for combining it with benthic.

```{r}
# Load phylogenetic trees
# Read in all 500 trees
sharkTrees <- read.nexus('Data/SharkData/PhyloTrees.nex')

# Set the seed
set.seed(125)

# Sample out a single tree
singleTree <- sample(sharkTrees, 1)[[1]]

# Create correlation structure
corBM <- corBrownian(phy = singleTree, form = ~ Genus_species)

# Model 1: original three categories
OGwaterCategories <- gls(CFAR ~ WaterColumnPosition, 
            data = categoricalData, 
            correlation = corBM,
            method = "ML")

summary(OGwaterCategories)

# Model 2: benthopelagic > pelagic
pelagicPosition <- categoricalData %>%
  mutate(WaterColumnPosition = case_when(
   str_detect(WaterColumnPosition, "Pelagic") ~ "Pelagic",
   WaterColumnPosition == "Benthic" ~ "Benthic",
   WaterColumnPosition == "Benthopelagic" ~ "Pelagic"))

ToPelagic <- gls(CFAR ~ WaterColumnPosition, 
            data = pelagicPosition, 
            correlation = corBM,
            method = "ML")

summary(ToPelagic)

# Model 3: benthopelagic > benthic
benthicPosition <- categoricalData %>%
  mutate(WaterColumnPosition = case_when(
   str_detect(WaterColumnPosition, "Pelagic") ~ "Pelagic",
   WaterColumnPosition == "Benthic" ~ "Benthic",
   WaterColumnPosition == "Benthopelagic" ~ "Benthic"))

ToBenthic <- gls(CFAR ~ WaterColumnPosition, 
            data = benthicPosition, 
            correlation = corBM,
            method = "ML")

summary(ToBenthic)

# AIC comparison
compareWaterColumnAIC <- data.frame(AIC(OGwaterCategories, ToPelagic, ToBenthic))

compareWaterColumnAIC

# Save AIC data frame as a csv
write.csv(compareWaterColumnAIC, "Supplementary_Material/Results/TableS2_WaterPosition.csv",
          row.names = TRUE)
```

## Step 4: Compare coastal, deepwater, and pelagic habitat use

The second approach is to compare CFAR values for the three habitat use levels:
* Coastal
* Deepwater
* Oceanic (Pelagic)

Note: to avoid confusion with the water column position "Pelagic", we rename "Pelagic" to "Oceanic" for habitat use.

We'll first visualize the spread of CFAR by habitat.

```{r}
categoricalData <- categoricalData %>%
  mutate(PrimaryHabitat = case_when(
   str_detect(PrimaryHabitat, "Pelagic") ~ "Oceanic",
   PrimaryHabitat == "Deepwater" ~ "Deepwater",
   PrimaryHabitat == "Coastal" ~ "Coastal"))

habitat_order <- c("Deepwater", "Coastal", "Oceanic")

FigureS3 <- ggplot() +
  geom_jitter(data = categoricalData, 
             aes(x = factor(PrimaryHabitat, levels = habitat_order), y = CFAR,
                 color = PrimaryHabitat), size = 2) +
  labs(x = element_blank(),
       y = "Caudal Fin Aspect Ratio (CFAR)") +
  scale_colour_manual(values = c("Oceanic" = "#97a55d", 
                               "Coastal" = "#6a7132",
                               "Deepwater" = "#3e4415")) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.y = element_text(size = 12),
        legend.position = "none")

FigureS3

# Save plot as a png
ggsave("Supplementary_Material/Results/FigS3.tiff",
       plot = FigureS3, units = "px",
       device = "tiff",
       width = 1200, height = 1300, dpi = 300)
```

Now we'll statistically evaluate which category deepwater should be binned with.

Biologically, we expect support for combining it with coastal.

For clarity, we're also going to rename "Pelagic" to "Oceanic".

```{r}
# Model 1: original three categories
OGhabitatCategories <- gls(CFAR ~ PrimaryHabitat, 
            data = categoricalData, 
            correlation = corBM,
            method = "ML")

summary(OGhabitatCategories)

# Model 2: deepwater > coastal
coastalUse <- categoricalData %>%
  mutate(PrimaryHabitat = case_when(
   str_detect(PrimaryHabitat, "Oceanic") ~ "Oceanic",
   PrimaryHabitat == "Deepwater" ~ "Coastal",
   PrimaryHabitat == "Coastal" ~ "Coastal"))

ToCoastal <- gls(CFAR ~ PrimaryHabitat, 
            data = coastalUse, 
            correlation = corBM,
            method = "ML")

summary(ToCoastal)

# Model 3: deepwater > oceanic
oceanicUse <- categoricalData %>%
  mutate(PrimaryHabitat = case_when(
   str_detect(PrimaryHabitat, "Oceanic") ~ "Oceanic",
   PrimaryHabitat == "Deepwater" ~ "Oceanic",
   PrimaryHabitat == "Coastal" ~ "Coastal"))

ToOceanic <- gls(CFAR ~ PrimaryHabitat, 
            data = oceanicUse, 
            correlation = corBM,
            method = "ML")

summary(ToOceanic)

# AIC comparison
compareHabitatAIC <- data.frame(AIC(OGhabitatCategories, ToCoastal, ToOceanic))

compareHabitatAIC

# Save AIC data frame as a csv
write.csv(compareHabitatAIC, "Supplementary_Material/Results/TableS3_HabitatUse.csv",
          row.names = TRUE)
```

## Step 5: Compare CFAR between the four combinations of habitat and water column position

Now sharks should either be:
* Oceanic X Pelagic
* Oceanic X Benthic
* Coastal X Pelagic
* Coastal X Benthic

We expect most sharks should be in Coastal X Benthic.

```{r}
catCondensed <- categoricalData %>%
  # Benthopelagic > Benthic
  mutate(WaterColumnPosition = case_when(
   str_detect(WaterColumnPosition, "Pelagic") ~ "Pelagic",
   WaterColumnPosition == "Benthic" ~ "Benthic",
   WaterColumnPosition == "Benthopelagic" ~ "Benthic")) %>%
  # Deepwater > Coastal
  mutate(PrimaryHabitat = case_when(
   str_detect(PrimaryHabitat, "Oceanic") ~ "Oceanic",
   PrimaryHabitat == "Deepwater" ~ "Coastal",
   PrimaryHabitat == "Coastal" ~ "Coastal"))

FigureS4 <- ggplot(catCondensed, 
                              aes(x = PrimaryHabitat, y = WaterColumnPosition, 
                                  color = CFAR)) +
  geom_jitter(size = 3, alpha = 0.8, width = 0.3, height = 0.3) +
  scale_color_scico(palette = "corkO") +
  theme_classic() +
  labs(x = "Habitat Use",
       y = "Water Column Position",
       color = "CFAR") +
  theme(legend.position = "inside",
        legend.position.inside = c(0.92, 0.2))

FigureS4

# Save plot
ggsave("Supplementary_Material/Results/FigS4.tiff",
       plot = FigureS3, units = "px",
       device = "tiff",
       width = 1500, height = 1500, dpi = 300)
```

We see that most of the data falls into either Coastal X Benthic or Oceanic X Pelagic.