---
title: "3_phylopath"
output: html_document
---

## Purpose

To test data consistency and get coefficients using the 'phylopath' package.

Phylopath tests d-separation while accounting for phylogeny.

This supplemental analysis demonstrates why average dissolved oxygen needed to be excluded.

### Step 1: Load libraries

Load the libraries used.

```{r}
# Load tidyverse for sanity
library(tidyverse)

# Load ape for phylogenetic tree manipulation
library(ape)

# Load phylopath for analysis
library(phylopath)

# Load car for VIF
library(car)

# Load packages for plotting
library(cowplot)
library(png)
library(grid)
```

### Step 2: Load data and create models to compare

```{r}
# Load shark data
sharkData <- read.csv("Data/SharkData/TraitData.csv") %>%
  # See 1_Supplemental_CategoricalAnalysis for Benthopelagic > Benthic and Deepwater > Coastal
  # Benthopelagic > Benthic
  mutate(WaterColumnPosition = case_when(
   str_detect(WaterColumnPosition, "Pelagic") ~ "Pelagic",
   WaterColumnPosition == "Benthic" ~ "Benthic",
   WaterColumnPosition == "Benthopelagic" ~ "Benthic")) %>%
  # Deepwater > Coastal
  mutate(PrimaryHabitat = case_when(
   str_detect(PrimaryHabitat, "Pelagic") ~ "Oceanic",
   PrimaryHabitat == "Deepwater" ~ "Coastal",
   PrimaryHabitat == "Coastal" ~ "Coastal"))

# Load climate data to append oxygen
oxygenData <- read.csv("Supplementary_Material/Data/SharkData/SpeciesClimateData_251217.csv") %>%
  # Use filter to remove chimaeras and rays
  filter(order_name != 'MYLIOBATIFORMES', 
         order_name != 'RAJIFORMES',
         order_name != 'RHINOPRISTIFORMES',
         order_name != 'TORPEDINIFORMES') %>%
  mutate(Genus_species = str_replace_all(scientific_name, " ", "_")) %>%
  select(Genus_species, average_dissolved_oxygen_micromoles_per_kg) %>%
  distinct()

# Use left_join to add oxygen data to sharkData
allSharkData <- sharkData %>%
  left_join(oxygenData, by = "Genus_species")

# Rename columns for sanity
sharkDataAbrv <- allSharkData %>% 
  rename(GillHeight = ProportionalGillAverage,
         MaxBodySize = MaximumBodySize,
         AvgTemp = AvgTemperature,
         WaterColumn = WaterColumnPosition,
         HabitatUse = PrimaryHabitat,
         AvgDissolvedO2 = average_dissolved_oxygen_micromoles_per_kg)

# Convert WaterColumn and HabitatUse to a factor
sharkDataAbrv$WaterColumn <- as.factor(sharkDataAbrv$WaterColumn)
sharkDataAbrv$HabitatUse <- as.factor(sharkDataAbrv$HabitatUse)

# Log Maximum Body Size
sharkDataAbrv$MaxBodySize = log(sharkDataAbrv$MaxBodySize)

# Set variables to be scaled
vars_to_scale <- c("MaxBodySize", "GillHeight", "AvgTemp", "AvgDissolvedO2")

# Center and scale continuous variables
sharkDataScaled <- sharkDataAbrv %>%
  mutate(across(all_of(vars_to_scale), scale)) %>%
  # Log CFAR due to Gamma log-linked distribution
  mutate(CFAR = log(CFAR))

# Define competing model sets
models <- define_model_set(
  # WaterColumn as a consequence of HabitatUse
  DO_One   = c(WaterColumn ~ HabitatUse), 
  # HabitatUse as a consequence of WaterColumn
  DO_Two   = c(HabitatUse ~ WaterColumn),
  # Relationships all models have in common
  .common = c(MaxBodySize ~ AvgTemp,
              HabitatUse ~ AvgTemp,
              MaxBodySize ~ WaterColumn,
              AvgDissolvedO2 ~ AvgTemp,
              GillHeight ~ MaxBodySize,
              GillHeight ~ AvgDissolvedO2,
              GillHeight ~ WaterColumn,
              CFAR ~ WaterColumn,
              CFAR ~ GillHeight,
              CFAR ~ MaxBodySize,
              CFAR ~ AvgDissolvedO2,
              CFAR ~ HabitatUse)
  )

# Create data frame for label positions
positions <- data.frame(
  name = c('AvgTemp', 
           'HabitatUse','MaxBodySize', 
           'WaterColumn', 'GillHeight', 'AvgDissolvedO2', 
           'CFAR'),
  x = c(1.7, 
        1.3, 2, 
        1.2, 1.9, 1.4,
        2),
  y = c(4, 
        3.5, 3.4, 
        1.7, 2.1, 2.8,
        1)
)

plot_model_set(models, manual_layout = positions, edge_width = 0.75,
               box_x = 35, box_y = 10, curvature = 0.1)
```

### Step 4: Load tree data

```{r}
# Read in all 500 trees
FullTree <- read.nexus('Data/SharkData/PhyloTrees.nex')

# Set the seed
set.seed(125)

# Sample out a single tree
SampleFullTree <- sample(FullTree, 1)[[1]]

# Check if tips are binary
is.binary(SampleFullTree)

# Check tip labels
SampleFullTree$tip.label

# Check length of df = length of tree (should print TRUE)
length(sharkData$Genus_species) == length(SampleFullTree$tip.label)

# Set species names as row names
rownames(sharkDataScaled) <- sharkDataScaled$Genus_species
```

### Step 5: Evaluate the d-separation implications of DAG

```{r}
# Run phylo_path on the model sets
result <- phylo_path(models, data = sharkDataScaled, 
                     tree = SampleFullTree, model = "lambda", 
                     method = "logistic_MPLE")
# View the results
result

# Summary of results
s <- summary(result)

# View the summary
s

# Export dissolved oxygen results
dissolvedOxygen <- as.data.frame(s)

write.csv(dissolvedOxygen,
          "Supplementary_Material/Results/TableS8_DOmodel.csv",
          row.names = FALSE)

# Plot the relative importance of both models
importancePlot <- plot(s)
importancePlot

dissolvedOxygen <- dissolvedOxygen %>%
  mutate(
    LegendGroup = if_else(delta_CICc <= 2,
                          "Within 2 CICc",
                          "Not within 2 CICc")
  )

# Basically recreate the default plot, but pretty
dO2Plot <- ggplot(dissolvedOxygen, 
                        aes(x = reorder(model, p), y = p,
                            fill = LegendGroup)) +
  geom_col(color = 'black') +
  geom_text(aes(label = round(p, 3)),
            hjust = 1,                
            size = 3,
            color = "white") +
  coord_flip() +
  scale_fill_manual(
    values = c("Within 2 CICc" = "#c97373",
               "Not within 2 CICc" = "#5e5e5e")) +
  labs(x = NULL, y = NULL) +
  theme_classic() +
  theme(axis.text = element_text(size = 5),
    legend.text = element_text(size = 7),
    legend.title = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.2),
    legend.key.size = unit(0.5, 'cm'))

dO2Plot
```

Create Figure S6.

```{r}
# Upload nice looking DAG
FigS6a <- readPNG("Supplementary_Material/Plot_Inserts/FigS6a.png")

# Convert to grob
panelA <- rasterGrob(FigS6a)

# Combine
FigureS6 <- plot_grid(
  panelA, dO2Plot, rel_widths = c(0.5, 1.1),
  ncol = 2, labels = c("a", "b"),
  label_size = 8
)

# Save model comparison plot as a png
ggsave("Supplementary_Material/Results/FigS6.tiff",
       plot = FigureS6, units = "px",
       device = "tiff", bg = "white",
       width = 1200, height = 600, dpi = 300)
```


```{r}
# Manually inspect d_sep results
result$d_sep$DO_One
result$d_sep$DO_Two
```

AvgDissolvedO2 is a recurring problem child in the d_sep results.

We know that dissolved oxygen is strongly correlated with temperature because of chemistry.

Therefore, we can investigate what happens if we drop average dissolved oxygen.

```{r}
sharkDataScaled1 <- sharkDataScaled %>%
  select(-AvgDissolvedO2)

# Define competing model sets
modelsNoO2 <- define_model_set(
  # WaterColumn as a consequence of HabitatUse
  Hab_One   = c(WaterColumn ~ HabitatUse),
  # HabitatUse as a consequence of WaterColumn
  Hab_Two   = c(HabitatUse ~ WaterColumn),
  # Relationships all models have in common
  .common = c(MaxBodySize ~ AvgTemp,
              HabitatUse ~ AvgTemp,
              MaxBodySize ~ WaterColumn,
              GillHeight ~ MaxBodySize,
              GillHeight ~ WaterColumn,
              CFAR ~ WaterColumn,
              CFAR ~ GillHeight,
              CFAR ~ MaxBodySize,
              CFAR ~ HabitatUse)
  )

# Create data frame for label positions
positionsNoO2 <- data.frame(
  name = c('AvgTemp', 
           'HabitatUse','MaxBodySize', 
           'WaterColumn', 'GillHeight', 
           'CFAR'),
  x = c(1.7, 
        1.3, 2.1, 
        1.5, 1.4,
        2),
  y = c(4, 
        3.5, 3.4, 
        2.3, 1.5,
        1)
)

plot_model_set(modelsNoO2, manual_layout = positionsNoO2, edge_width = 0.75,
               box_x = 30, box_y = 7, curvature = 0.1)

# Run phylo_path on the model sets
resultNoO2 <- phylo_path(modelsNoO2, data = sharkDataScaled1, 
                     tree = SampleFullTree, model = "lambda", 
                     method = "logistic_MPLE")
# View the results
resultNoO2

# Summary of results
sO2 <- summary(resultNoO2)

# View the summary
sO2

# Export habitat use results
habitatUse <- as.data.frame(sO2)

write.csv(habitatUse,
          "Supplementary_Material/Results/TableS9_HabitatModel.csv",
          row.names = FALSE)

# Plot the relative importance of all four models
importancePlotNoO2 <- plot(sO2)
importancePlotNoO2

habitatUse <- habitatUse %>%
  mutate(
    LegendGroup = if_else(delta_CICc <= 2,
                          "Within 2 CICc",
                          "Not within 2 CICc")
  )

# Basically recreate the default plot, but pretty
habitatPlot <- ggplot(habitatUse, 
                        aes(x = reorder(model, p), y = p,
                            fill = LegendGroup)) +
  geom_col(color = 'black') +
  geom_text(aes(label = round(p, 3)),
            hjust = 1,                
            size = 3,
            color = "white") +
  coord_flip() +
  scale_fill_manual(
    values = c("Within 2 CICc" = "#c97373",
               "Not within 2 CICc" = "#5e5e5e")) +
  labs(x = NULL, y = NULL) +
  theme_classic() +
  theme(axis.text = element_text(size = 5),
    legend.text = element_text(size = 7),
    legend.title = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.2),
    legend.key.size = unit(0.5, 'cm'))

habitatPlot
```

```{r}
# Manually inspect d_sep results
resultNoO2$d_sep$Hab_One
resultNoO2$d_sep$Hab_Two
```

```{r}
# Extract the best model
bestNoO2 <- best(resultNoO2)

best_positions <- data.frame(
  name = c('AvgTemp', 
           'HabitatUse',
           'MaxBodySize', 
           'WaterColumn', 
           'GillHeight', 
           'CFAR'),
  x = c(5, 
        2.5, 
        8.2, 
        2.2, 
        8.5,
        5),
  y = c(6, 
        5, 
        4, 
        3, 
        2,
        0.5)
)

# Plot the best model
bestModel <- plot(bestNoO2, manual_layout = best_positions, 
                  edge_width = 0.75, text_size = 3,
                  box_x = 20, box_y = 5, curvature = 0.02) +
             ggplot2::xlim(0,11) +
             ggplot2::ylim(0.2, 6.05) +
             ggplot2::theme(legend.position = "none")

# View the best model
bestModel
```

Create Figure S7.

```{r}
# Upload nice looking DAG
FigS7a <- readPNG("Supplementary_Material/Plot_Inserts/FigS7a.png")

# Convert to grob
panel7A <- rasterGrob(FigS7a)

# Create left column
leftCol <- plot_grid(
  panel7A,
  habitatPlot,
  rel_widths = c(1, 0.5),
  ncol = 1, labels = c("a", "b"),
  label_size = 8
)

# Combine with model
FigureS7 <- plot_grid(
  leftCol, bestModel, rel_widths = c(1, 1.1),
  ncol = 2, labels = c("", "c"),
  label_size = 8
)

# Save best model plot as a png
ggsave("Supplementary_Material/Results/FigS7.tiff",
       plot = FigureS7, units = "px",
       device = "tiff", bg = "white",
       width = 1500, height = 2000, dpi = 300)
```

By removing dissolved oxygen, two models pass the d-separation test.

HabitatUse is a recurring problem, and the coefficient between 
Water Column Position > HabitatUse is orders of magnitude larger than any others.

This supports WaterColumn and HabitatUse are very correlated
(See 1_Supplemental_CategoricalAnalysis for more).

Therefore, we drop HabitatUse because WaterColumn is more ecologically relevant.
This is also supported by the path coefficient between HabitatUse and CFAR (0.01) versus 
the path coefficient between WaterColumn and CFAR (0.5).
A species is unlikely to "become oceanic" if they are "confined to the benthos", if we assume
that the ancestral state was benthic and coastal (Compagno 1990; Sternes et al. 2024).