---
title: "1_dispersalCFAR"
output: html_document
---

## Purpose

To demonstrate the utility of CFAR as a metric for dispersal ability:
* CFAR vs. swimming speed (from Iliou et al. 2023: https://doi.org/10.1098/rsos.231127)
* CFAR vs. geographic range

### Step 1: Load libraries

Load the libraries used.

```{r}
# Load tidyverse for dplyr and ggplot2
library(tidyverse)

# Load ape, geiger, and phytools for dealing with phylogenetic trees
library(ape)
library(geiger)
library(phytools)

# Load brms for analysis
library(brms)

# Load tidybayes for analyzing brms
library(tidybayes)

# Load packages  for plotting
library(scales)
library(cowplot)
library(png)
library(grid)
library(gridExtra)
```

## Step 2: Load data and phylogenetic trees

```{r}
## Swimming Speed Data ##
# Downloaded and modified from Iliou et al. 2023 (doi:10.6084/m9.figshare.c.6858294)
jacobyData <- read.csv("Data/SharkData/Iliou_SpeedData.csv") %>%
  mutate(Genus_species = str_replace(Species, ".*,\\s*", "")) %>%
  mutate(Genus_species = str_replace_all(Genus_species, " ", "_")) %>%
  select(Genus_species, speed, CFAR) %>%
  # Fix species name typos
  mutate(Genus_species = case_when(
    Genus_species == "Galeocerdo_curvier" ~ "Galeocerdo_cuvier",
    Genus_species == "Carcharhinus_obscures" ~ "Carcharhinus_obscurus",
    Genus_species == "Rhinocodon_typus" ~ "Rhincodon_typus",
    TRUE ~ Genus_species
  ))

## Trait Data ##
sharkData <- read.csv("Data/SharkData/TraitData.csv") %>%
  # See 1_Supplemental_CategoricalAnalysis for Benthopelagic > Benthic and Deepwater > Coastal
  # Benthopelagic > Benthic
  mutate(WaterColumnPosition = case_when(
    str_detect(WaterColumnPosition, "Pelagic") ~ "Pelagic",
    WaterColumnPosition == "Benthic" ~ "Benthic",
    WaterColumnPosition == "Benthopelagic" ~ "Benthic")) %>%
  left_join(select(read.csv("Data/SharkData/IUCNrangeData.csv"), 
                   "Genus_species", "area_sqkm"),
            by = "Genus_species")

## Phylogenetic Tree ##
# Read in all 500 trees
# There are different species in the trait data and swimming speed data so we need two "sets" of trees
traitTrees <- read.nexus("Data/SharkData/PhyloTrees.nex")
swimTrees <- traitTrees
```

## Step 3: Analyze swimming speed

Since the swimming speed data has way fewer species than the trait data, we need to trim the phylogenetic trees.

```{r}
## Tree trimming for swimming speed data ##
# Set the seed for reproducibility
set.seed(123)

# Sample out a single tree
swimTree <- sample(swimTrees, 1)[[1]]

# View which species are a mistmatch between the tree and data (should return Squalus_lalannei)
diffNamesSwim <- name.check(swimTree, jacobyData, jacobyData$Genus_species)

# Keep only tips that are present in the data
swimTreeDrop <- drop.tip(swimTree,
                        setdiff(swimTree$tip.label, 
                                jacobyData$Genus_species))

# Check names between tree and data
swimDiffNamesCheck <- name.check(swimTreeDrop, jacobyData, 
                                 jacobyData$Genus_species)

# Drop Cetorhinus maximus and Mustelus antarcticus from data
jacobyData <- jacobyData %>%
  filter(Genus_species != "Cetorhinus_maximus",
         Genus_species != "Mustelus_antarcticus")

# Check names again (should return "OK")
swimDiffNamesCheck <- name.check(swimTreeDrop, jacobyData, jacobyData$Genus_species)

# Apply dropped species to all trees
swimTreesTrim <- lapply(swimTrees, function(tr) {
  drop.tip(tr, setdiff(tr$tip.label, jacobyData$Genus_species))
})

# Reassign class so augmentedTreesTrim is recognized as a multiPhylo object
class(swimTreesTrim) <- "multiPhylo"

# Confirm all trees now match the data (should return TRUE)
all(sapply(swimTreesTrim, function(tr) {
  length(setdiff(jacobyData$Genus_species, tr$tip.label)) == 0
}))

# Sample out a single tree
swimTreeB <- sample(swimTreesTrim, 1)[[1]]

# Double-check that names match (should return "OK")
swimDiffNames2 <- name.check(swimTreeB, jacobyData, jacobyData$Genus_species)
```

Now analyze swimming speed with brms.

```{r}
## Anthony's Code ##
# Scale CFAR
jacobyDataScaled <- jacobyData
jacobyDataScaled$CFAR <- scale(jacobyData$CFAR)

# Create covariance matrix
A <- ape::vcv(swimTreeB)
  
# Swimming Speed ~ CFAR + Phylogeny
swim_prior <- get_prior(speed ~ CFAR + (1|gr(Genus_species, cov = A)),
                      data = jacobyDataScaled,
                      data2 = list(A = A),
                      family = gaussian())
# Set priors 
priors_swim <- c(
  # Mean speed around 0.3
  prior(normal(0.3, 0.1), class = "Intercept"),  
  prior(normal(0, 0.2), class = "b"),    
  # Residual SD likely ~0.1–0.3
  prior(exponential(5), class = "sigma"),
  # Random effect SDs (including phylogenetic term)
  prior(exponential(2), class = "sd")
)

swimSpeed <- brm(speed ~ CFAR + (1|gr(Genus_species, cov = A)),
              data = jacobyData, 
              data2 = list(A = A),
              family = gaussian(),
              prior = priors_swim,
              # Increase adapt_delta due to small sample size
              control = list(adapt_delta = 0.97),
              warmup = 1000, iter = 3000,
              cores = 2,
              save_pars = save_pars(all = TRUE))

summary(swimSpeed)

# Save brms object
saveRDS(swimSpeed, "Results/brmsOutputs/brmsSwimSpeed.rds")
```

Optional: load brms object if revisiting script.

```{r}
# Load brms objects using readRDS
swimSpeed <- readRDS("Results/brmsOutputs/brmsSwimSpeed.rds")
```

Plot CFAR vs. swim speed

```{r}
# Tibble of results
CFAR_swim <- tibble(
  CFAR = seq(
    min(jacobyData$CFAR, na.rm = TRUE),
    max(jacobyData$CFAR, na.rm = TRUE),
    # Smooth the line
    length.out = 1000)
  )

# Expected values
swim_epred_draws <- add_epred_draws(CFAR_swim, swimSpeed, ndraws = 500, 
                                    re_formula = NA)

# Append habitat data
jacobyDataHabitat <- jacobyData %>%
  left_join(select(sharkData, Genus_species, PrimaryHabitat), by = "Genus_species")

# Plot
speedPlot <- ggplot(swim_epred_draws, aes(x = CFAR, y = .epred)) +
  stat_lineribbon(.width = 0, color = "#dbb6bd", alpha = 0.5) +  
  geom_line(aes(group = .draw), color = "#dbb6bd", linewidth = 0.1, alpha = 0.1) +
  geom_point(data = jacobyDataHabitat, aes(x = CFAR, y = speed,
                                           color = PrimaryHabitat),
             alpha = 0.5, size = 1) +
  ylim(0,1.5) +
  scale_color_manual(values = c("#b37c86",
                                "#b37c86",
                                "#528598"),
                     labels = c("Deepwater",
                                "Coastal",
                                "Pelagic")) +
  labs(#title = "B",
       x = "Caudal Fin Aspect Ratio", 
       y = bquote('Swimming speed ('*m%.%s^{-1}*')')) +
  theme_classic() +
  theme(legend.position = "none",
        # plot.title.position = "plot",
        # plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 5),
        axis.title.x = element_text(size = 6),
        axis.text.y = element_text(size = 5),
        axis.title.y = element_text(size = 6),
        plot.margin = unit(c(0.02, 0.17, 0, 0), "cm"))

speedPlot
```

## Step 4: Analyze geographic range

```{r}
# Set the seed for reproducibility
set.seed(125)

# Sample out a single tree
singleTree <- sample(traitTrees, 1)[[1]]

# Check if/where the tree and trait data differ (should return "OK")
nameDrops <- name.check(singleTree, sharkData, sharkData$Genus_species)

# Check length of df = length of tree (should return TRUE)
length(sharkData$Genus_species) == length(singleTree$tip.label)

# Reorder data to match tree just in case
target <- singleTree$tip.label

# Reorder sharkData rows in the same order as singleTree$tip.label
sharkData <- sharkData[match(target, sharkData$Genus_species), ]

# Rename rows to match tip labels
rownames(sharkData) <- sharkData$Genus_species

# Set variables to be scaled
vars_to_scale <- c("MaximumBodySize", "ProportionalGillAverage", 
                   "AvgTemperature", "CFAR")

# Center and scale continuous variables (vars_to_scale)
sharkDataScaled <- sharkData %>%
  mutate(across(all_of(vars_to_scale), scale))
```

Run brms on geographic range vs CFAR

```{r}
# Create variance covariance matrix
B <- ape::vcv(singleTree)

## CFAR ~ Geographic Range + Phylogeny
geoRange <- brm(area_sqkm ~ CFAR + (1|gr(Genus_species, cov = B)),
          data = sharkDataScaled,
          data2 = list(B = B),
          family = Gamma(link = "log"),
          warmup = 1000, iter = 3000,
          cores = 2,
          save_pars = save_pars(all = TRUE))

summary(geoRange)

# Save brms object
saveRDS(geoRange, "Results/brmsOutputs/brmsGeographicRange.rds")
```

## Optional: Load brms object if revisiting script

```{r}
# Load brms objects using readRDS
geoRange <- readRDS("Results/brmsOutputs/brmsGeographicRange.rds")
```

Plot CFAR vs. Geographic Range

```{r}
# Tibble of results
CFAR_geo <- tibble(
  CFAR = seq(
    min(sharkData$CFAR, na.rm = TRUE),
    max(sharkData$CFAR, na.rm = TRUE),
    # Smooth the line
    length.out = 1000)
  )

# Expected values
geo_epred_draws <- add_epred_draws(CFAR_geo, geoRange, ndraws = 500, 
                                    re_formula = NA)

rangePlot <- ggplot(geo_epred_draws, aes(x = CFAR, y = .epred)) +
  stat_lineribbon(.width = 0, color = "#775d8c", alpha = 0.5) +  
  geom_line(aes(group = .draw), color = "#775d8c", linewidth = 0.1, alpha = 0.1) +
  geom_point(data = sharkData, aes(x = CFAR, y = area_sqkm,
                                   color = PrimaryHabitat),
             alpha = 0.5, size = 1) +
  scale_y_log10(# Set breaks at specific orders of magnitude
    breaks = trans_breaks("log10", function(x) 10^x),
    # Format labels nicely as scientific notation
    labels = trans_format("log10", math_format(10^.x))) +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 5), 
                     labels = c(1, 2, 3, 4, 5)) +
  scale_color_manual(values = c("#2e193f",
                                "#2e193f",
                                "#528598"),
                     labels = c("Deepwater",
                                "Coastal",
                                "Pelagic")) +
  labs(# title = "C",
       x = "Caudal Fin Aspect Ratio", 
       y = bquote('Geographic Range ('*km^{2}*')')) +
  theme_classic() +
  theme(legend.position = "none",
        # plot.title.position = "plot",
        # plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 5),
        axis.title.x = element_text(size = 6),
        axis.text.y = element_text(size = 5),
        axis.title.y = element_text(size = 6))

rangePlot
```

Warning about 4 rows containing missing values are for:
*Aulohalaelurus kanakorum*
*Carcharhinus perezii*
*Cephaloscyllium zebrum*
*Chiloscyllium hasseltii*

Which are missing geographic range data.

Now, arrange plots and save.

```{r}
# Upload panel A (CFAR diagram)
panelA_image <- readPNG("Plot_Inserts/Figure1A.png")

panelA <- ggdraw() +
  draw_image(panelA_image, scale = 1)

panelA
```


```{r}
# Align plots by y-axis
aligned <- cowplot::align_plots(speedPlot, rangePlot, align = "hv")

# Create spacer
space <- nullGrob()

# Arrange plots
Figure1 <- plot_grid(
  panelA,
  space,
  aligned[[1]],
  aligned[[2]],
  ncol = 1,
  rel_heights = c(0.5, 0.02, 1, 1),
  labels = c("a", "", "b", "c"),
  label_size = 8
)

# Save as Fig 1
ggsave("Results/Figures/Fig1.tiff", 
       plot = Figure1, units = "px",
       device = "tiff",
       bg = "white",
       width = 700, height = 1400,
       dpi = 300)
```
