---
title: "2_phylogeneticAnalysis"
output: html_document
---

## Purpose

To visualize the phylogenetic pattern in CFAR and other traits.

### Step 1: Load libraries

Load the libraries used.

```{r}
# Load tidyverse for sanity
library(tidyverse)

# Load phytools, ape, and geiger for working with phylogenetic trees
library(phytools)
library(ape)
library(geiger)

# Load ggtree and ggtreeExtrafor plotting phylogenetic tree
library(ggtree)
library(ggtreeExtra)
library(ggnewscale)

# Load scico for color palettes
library(scico)

# Load phylopic for silhouettes
library(rphylopic)

# Load plotting libraries
library(cowplot)
library(grid)
library(gridExtra)
```

### Step 2: Load files

Load:
* Nexus file (PhyloTrees.nex) 
* Shark trait data (TraitData.csv)

```{r}
# Read the tree file
Trees <- read.nexus("Data/SharkData/PhyloTrees.nex")

# Use read.csv to load shark trait data
sharkData <- read.csv("Data/SharkData/TraitData.csv")

# Load Matley tracking data
trackData <- read.csv("Supplementary_Material/Data/SharkData/Matley_TaggedSpecies.csv") %>%
  # Mutate species names to match trait data
  mutate(Genus_species = str_replace_all(species_fishbase, " ", "_")) %>%
  select(Genus_species, n.tagged)
```

### Step 3: Sample a tree and make sure they match

```{r}
# Set seed for reproducability
set.seed(123)

# Sample a single tree from Trees
sharkTree <- sample(Trees, 1)[[1]]

# Set row names as species names
rownames(sharkData) <- sharkData$Genus_species

# Check species with setdiff (should return "character (0)")
setdiff(sharkTree$tip.label, sharkData$Genus_species)
```

### Step 4: Append tracking data

```{r}
sharkJoined <- sharkData %>%
  # Use left_join
  left_join(trackData, by = "Genus_species") %>%
  mutate(Tracked = if_else(is.na(n.tagged), "No","Yes"))
```

### Step 5: Get some higher level taxonomic information

```{r}
# Create a data frame of just the data we want
sharkTraits <- sharkJoined %>%
  select(Genus_species, Order, CFAR, MaximumBodySize, 
        ProportionalGillAverage, WaterColumnPosition, Tracked) %>%
# See 1_Supplemental_CategoricalAnalysis for Benthopelagic > Benthic and Deepwater > Coastal
  # Benthopelagic > Benthic
  mutate(WaterColumnPosition = case_when(
    str_detect(WaterColumnPosition, "Pelagic") ~ "Pelagic",
    WaterColumnPosition == "Benthic" ~ "Benthic",
    WaterColumnPosition == "Benthopelagic" ~ "Benthic"))
```

```{r}
# Create a named list of species per Order
order_groups <- split(sharkTraits$Genus_species, sharkTraits$Order)

# Get MRCA nodes
order_nodes <- sapply(order_groups, function(x) getMRCA(sharkTree, x))
order_nodes
```

### Step 6: Create mega phylogenetic tree plot

```{r}
# Plot simple tree before we go wild
phyloTree <- ggtree(sharkTree, linewidth = 0.3)

# Look at tree
phyloTree
```

```{r}
# Get vector of node IDs
nodes_vector <- as.numeric(order_nodes)

# Assign names to vector elements
names(nodes_vector) <- names(order_nodes)

# Group ancestral clades
grouped_sharkTree <- groupClade(sharkTree, .node = nodes_vector)

# Plot tree with colored Orders
phyloTree <- ggtree(grouped_sharkTree, aes(color = group), 
                    linewidth = 0.3) +
  scale_color_manual(values = c(
    "Carcharhiniformes"  = "#ba4848",
    "Lamniformes"        = "#f7c435",
    "Orectolobiformes"   = "#f0b6ad",
    "Heterodontiformes"  = "#0b5227",
    "Squaliformes"       = "#85a993",
    "Pristiophoriformes" = "#dc8864",
    "Echinorhiniformes"  = "#818b2e",
    "Hexanchiformes"     = "#c75a1b",
    "0"                  = "black")) +
  guides(color = "none") +
  xlim(0, 3)

phyloTree
```

Create function for extracing legends.

```{r}
# Function to extract legends
get_legend <- function(p) {
  g <- ggplotGrob(p)
  legend <- g$grobs[
    sapply(g$grobs, function(x) x$name) == "guide-box"
  ][[1]]
  return(legend)
}
```

CFAR tree.

```{r}
# Add CFAR bars
phyloTreeCFAR <- phyloTree +
  theme(legend.position = "none") +
  new_scale_fill() +
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = CFAR, fill = CFAR),
    orientation = "y",
    stat = "identity",
    pwidth = 0.1
  ) +
  scale_fill_scico(palette = "cork") +
  theme(legend.direction = "horizontal",
        legend.position = "inside",
        legend.position.inside = c(0.1, 1),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"),
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5, , size = 7),
        legend.text = element_text(size = 6),
        plot.margin = unit(c(1, 0, 0.1, 0), "cm"))

# Extract legend
CFARlegend <- get_legend(phyloTreeCFAR)

# See tree
phyloTreeCFAR
```

Maximum Body Size tree.

```{r}
phyloTreeSize <- phyloTree +
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = log(MaximumBodySize), 
                  fill = log(MaximumBodySize)),
    orientation = "y",
    stat = "identity",
    pwidth = 0.1,
    inherit.aes = FALSE
  ) +
  scale_fill_scico(palette = "bam") +
  labs(fill = "Body Size") +
  theme(legend.direction = "horizontal",
        legend.position = "inside",
        legend.position.inside = c(0.1, 1),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"),
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5, size = 7),
        legend.text = element_text(size = 6),
        plot.margin = unit(c(1, 0, 0.1, 0), "cm"))

# Extract legend
sizeLegend <- get_legend(phyloTreeSize)

# See tree
phyloTreeSize
```

Gill Slit tree.

```{r}
phyloTreeGills <- phyloTree +
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = ProportionalGillAverage, fill = ProportionalGillAverage),
    orientation = "y",
    stat = "identity",
    pwidth = 0.1,
    inherit.aes = FALSE
  ) +
  scale_fill_scico(palette = "grayC",
                   breaks = c(0.02, 0.05, 0.08)) +
  labs(fill = "Gill Slit Height") +
  theme(legend.direction = "horizontal",
        legend.position = "inside",
        legend.position.inside = c(0.1, 1),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"),
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5, size = 7),
        legend.text = element_text(size = 6),
        plot.margin = unit(c(1, 0, 0.1, 0), "cm"))

# Extract legend
gillLegend <- get_legend(phyloTreeGills)

# See tree
phyloTreeGills
```

Water Column Position Tree.

```{r}
phyloTreePosition <- phyloTree +
  geom_fruit(
    data = sharkTraits,
    geom = geom_tile,
    mapping = aes(y = Genus_species, x = 0, 
                  fill = WaterColumnPosition),
    stat = "identity",
    width = 0.01,
    pwidth = 0.1
  ) +
  scale_fill_manual(values = c("Pelagic" = "#82e1da", 
                               "Benthic" = "#103f4d")) +
  labs(fill = "Position") +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(hjust = 0.5, size = 7),
        # legend.direction = "horizontal",
        legend.position = "inside",
        legend.title.position = "top",
        legend.position.inside = c(0.1, 0.8),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"))

# Extract legend
posLegend <- get_legend(phyloTreePosition)

# See tree
phyloTreePosition
```

Tracking data tree.

```{r}
phyloTreeTag <- phyloTree +
  geom_fruit(
    data = sharkTraits,
    geom = geom_tile,
    mapping = aes(y = Genus_species, x = 0, 
                  fill = Tracked),
    stat = "identity",
    width = 0.01,
    pwidth = 0.1
  ) +
  scale_fill_manual(values = c("Yes" = "black",
                               "No" = "white")) +
  labs(fill = "Tagged") +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(hjust = 0.5, size = 7),
        # legend.direction = "horizontal",
        legend.position = "inside",
        legend.title.position = "top",
        legend.position.inside = c(0.1, 0.8),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"))

# Extract legend
tagLegend <- get_legend(phyloTreeTag)

# See tree
phyloTreeTag
```

Create the mega tree.

```{r}
# Build tree without legends
phyloTreeFull <- phyloTree +
  # CFAR
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = CFAR, fill = CFAR),
    orientation = "y",
    stat = "identity",
    pwidth = 0.3,
    offset = 0.01) +
  scale_fill_scico(palette = "cork") +
  new_scale_fill() +
  # Maximum body size
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = log(MaximumBodySize), 
                  fill = log(MaximumBodySize)),
    orientation = "y",
    stat = "identity",
    pwidth = 0.3,
  ) +
  scale_fill_scico(palette = "bam") +
  new_scale_fill() +
  # Gill slit height
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = ProportionalGillAverage, 
                  fill = ProportionalGillAverage),
    orientation = "y",
    stat = "identity",
    pwidth = 0.3
  ) +
  scale_fill_scico(palette = "grayC") +
  new_scale_fill() +
  # Water column position
  geom_fruit(
    data = sharkTraits,
    geom = geom_tile,
    mapping = aes(y = Genus_species, x = 0,
                  fill = WaterColumnPosition),
    stat = "identity",
    width = 0.02,
    pwidth = 0.07,
    offset = 0.075
  ) +
  scale_fill_manual(values = c("Pelagic" = "#82e1da",
                               "Benthic" = "#103f4d")) +
  new_scale_fill() +
  # Tracking status
    geom_fruit(
    data = sharkTraits,
    geom = geom_tile,
    mapping = aes(y = Genus_species, x = 0,
                  fill = Tracked),
    stat = "identity",
    width = 0.02,
    pwidth = 0.07,
    offset = 0.271
  ) +
  scale_fill_manual(values = c("Yes" = "black",
                               "No" = "white")) +
  new_scale_fill() +
  labs(fill = "") +
  theme(legend.position = "none",
        margins = unit(c(1.1, 0.5, 0.1, 0),"cm"))

# Add legends in custom position
phyloTreeLegends <- ggdraw() +
  draw_plot(phyloTreeFull) +
  draw_plot(tagLegend, x = 0.640, y = 0.175) +
  draw_plot(posLegend, x = 0.572, y = 0.175) +
  draw_plot(gillLegend, x = 0.480, y = -0.027) +
  draw_plot(sizeLegend, x = 0.381, y = -0.027) +
  draw_plot(CFARlegend, x = 0.285, y = -0.027)

phyloTreeLegends
```

Add Order labels.

```{r}
phyloTreeLabel <- phyloTreeLegends + 
  annotate("text", x = 0.85, y = 0.94, label = "Carcharhiniformes", 
           size = 2.5, color = "#ba4848") +
  annotate("text", x = 0.85, y = 0.4, label = "Lamniformes", 
           size = 2.5, color = "#f7c435") +
  annotate("text", x = 0.85, y = 0.35, label = "Orectolobiformes", 
           size = 2.5, color = "#f0b6ad") +
  annotate("text", x = 0.85, y = 0.28, label = "Heterodontiformes", 
           size = 2.5, color = "#0b5227") +
  annotate("text", x = 0.85, y = 0.2, label = "Squaliformes", 
           size = 2.5, color = "#85a993") +
  annotate("text", x = 0.85, y = 0.03, label = "Pristiophoriformes", 
           size = 2.5, color = "#dc8864") +
  annotate("text", x = 0.85, y = 0.02, label = "Echinorhiniformes", 
           size = 2.5, color = "#818b2e") +
  annotate("text", x = 0.85, y = 0.01, label = "Hexanchiformes", 
           size = 2.5, color = "#c75a1b")

phyloTreeLabel
```

Add silhouettes.

```{r}
# Get images from rphylopic
carIMG <- get_phylopic(uuid = "07eeaf0e-0284-4c1d-8415-ddd65a736232")
lamIMG <- get_phylopic(uuid = "a807f354-fd7c-4547-a7e8-573f7ca75e63")
oreIMG <- get_phylopic(uuid = "9a1c5e03-89e1-433b-b702-c0b945fd51fa")
hetIMG <- get_phylopic(uuid = "28a9e2ea-2221-4af7-97e4-61d130348ae5")
squIMG <- get_phylopic(uuid = "c786f963-7e91-4ecc-8351-400a0ecfdfcc")

# Add to tree
phyloTreeFinal <- phyloTreeLabel +
  add_phylopic(x = 0.85, y = 0.91, img = carIMG, height = 0.035,
               fill = "#ba4848") +
  add_phylopic(x = 0.85, y = 0.375, img = lamIMG, height = 0.035,
               fill = "#f7c435") +
  add_phylopic(x = 0.85, y = 0.335, img = oreIMG, height = 0.015,
               fill = "#f0b6ad") +
  add_phylopic(x = 0.85, y = 0.26, img = hetIMG, height = 0.025,
               fill = "#0b5227") +
  add_phylopic(x = 0.85, y = 0.18, img = squIMG, height = 0.021,
               fill = "#85a993")

phyloTreeFinal
```

### Step 8: Create and save Figure 2

```{r}
# Create a nullGrob to act as a spacer
spacer <- nullGrob()

# Combine plots
Figure2 <- arrangeGrob(
  spacer,
  phyloTreeFinal,
  ncol = 1,
  heights = c(0.01, 1)
)

# Save final figure
ggsave("Results/Figures/Fig2.tiff", Figure2, units = "px",
       device = "tiff", bg = "white",
       width = 2000, height = 2450, dpi = 300)
```
