---
title: "2_phylogeneticAnalysis"
output: html_document
---

## Purpose

To visualize the phylogenetic pattern in CFAR and other traits.

### Step 1: Load libraries

Load the libraries used.

```{r}
# Load tidyverse for sanity
library(tidyverse)

# Load phytools, ape, and geiger for working with phylogenetic trees
library(phytools)
library(ape)
library(geiger)

# Load ggtree and ggtreeExtrafor plotting phylogenetic tree
library(ggtree)
library(ggtreeExtra)
library(ggnewscale)

# Load scico 
library(scico)

# Load cowplot
library(cowplot)
```

### Step 2: Load files

Load:
* Nexus file (PhyloTrees.nex) 
* Shark trait data (TraitData.csv)

```{r}
# Read the tree file
Trees <- read.nexus("Data/SharkData/PhyloTrees.nex")

# Use read.csv to load shark trait data
sharkData <- read.csv("Data/SharkData/TraitData.csv")

# Load Matley tracking data
trackData <- read.csv("Supplementary_Material/Data/SharkData/Matley_TaggedSpecies.csv") %>%
  # Mutate species names to match trait data
  mutate(Genus_species = str_replace_all(species_fishbase, " ", "_")) %>%
  select(Genus_species, n.tagged)
```

### Step 3: Sample a tree and make sure they match

```{r}
# Set seed for reproducability
set.seed(123)

# Sample a single tree from Trees
sharkTree <- sample(Trees, 1)[[1]]

# Set row names as species names
rownames(sharkData) <- sharkData$Genus_species

# Check species with setdiff (should return "character (0)")
setdiff(sharkTree$tip.label, sharkData$Genus_species)
```

### Step 4: Append tracking data

```{r}
sharkJoined <- sharkData %>%
  # Use left_join
  left_join(trackData, by = "Genus_species") %>%
  mutate(Tracked = if_else(is.na(n.tagged), "No","Yes"))
```


### Step 5: Get some higher level taxonomic information

```{r}
# Create a data frame of just the data we want
sharkTraits <- sharkJoined %>%
  select(Genus_species, Order, CFAR, MaximumBodySize, WaterColumnPosition, Tracked) %>%
# See 1_Supplemental_CategoricalAnalysis for Benthopelagic > Benthic and Deepwater > Coastal
  # Benthopelagic > Benthic
  mutate(WaterColumnPosition = case_when(
    str_detect(WaterColumnPosition, "Pelagic") ~ "Pelagic",
    WaterColumnPosition == "Benthic" ~ "Benthic",
    WaterColumnPosition == "Benthopelagic" ~ "Benthic"))
```

```{r}
# Create a named list of species per Order
order_groups <- split(sharkTraits$Genus_species, sharkTraits$Order)

# Get MRCA nodes
order_nodes <- sapply(order_groups, function(x) getMRCA(sharkTree, x))
order_nodes

# Create colors for each Order
order_colors <- scico(n = length(order_nodes), palette = "lipari")
names(order_colors) <- names(order_nodes)
```

### Step 6: Plot phylogenetic tree

```{r}
# Plot the tree before we go wild
phyloTree <- ggtree(sharkTree)

# Look at tree
phyloTree
```

```{r}
# Get plotted data
tree_data <- phyloTree$data

# Compute y positions for all Orders
order_y <- sapply(names(order_nodes), function(o) {
  node_number <- order_nodes[o]
  tree_data$y[tree_data$node == node_number]
})

# Print node positions
order_y
```

CFAR tree.

```{r}
# Add colors to Orders
phyloTreeOrder <- ggtree(sharkTree, linewidth = 0.3)

for (o in names(order_nodes)) {
  phyloTreeOrder <- phyloTreeOrder +
    geom_hilight(node = order_nodes[o], fill = order_colors[o], alpha = 0.2)
}

# Function to extract legends
get_legend <- function(p) {
  g <- ggplotGrob(p)
  legend <- g$grobs[
    sapply(g$grobs, function(x) x$name) == "guide-box"
  ][[1]]
  return(legend)
}

# Add CFAR bars
phyloTreeCFAR <- phyloTreeOrder +
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = CFAR, fill = CFAR),
    orientation = "y",
    stat = "identity",
    pwidth = 0.1
  ) +
  scale_fill_scico(palette = "cork") +
  theme(legend.direction = "horizontal",
        legend.position = "inside",
        legend.position.inside = c(0.1, 1),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"),
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5, , size = 7),
        legend.text = element_text(size = 6),
        plot.margin = unit(c(1, 0, 0.1, 0), "cm"))

# Extract legend
CFARlegend <- get_legend(phyloTreeCFAR)

phyloTreeCFAR
```

Maximum Body Size tree.

```{r}
phyloTreeSize <- phyloTreeOrder +
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = log(MaximumBodySize), fill = log(MaximumBodySize)),
    orientation = "y",
    stat = "identity",
    pwidth = 0.1,
    inherit.aes = FALSE
  ) +
  scale_fill_scico(palette = "bam") +
  labs(fill = "Body Size") +
  theme(legend.direction = "horizontal",
        legend.position = "inside",
        legend.position.inside = c(0.1, 1),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"),
        legend.title.position = "top",
        legend.title = element_text(hjust = 0.5, size = 7),
        legend.text = element_text(size = 6),
        plot.margin = unit(c(1, 0, 0.1, 0), "cm"))

# Extract legend
sizeLegend <- get_legend(phyloTreeSize)

phyloTreeSize
```

Water Column Position Tree.

```{r}
phyloTreePosition <- phyloTreeOrder +
  geom_fruit(
    data = sharkTraits,
    geom = geom_tile,
    mapping = aes(y = Genus_species, x = 0, 
                  fill = WaterColumnPosition),
    stat = "identity",
    width = 0.01,
    pwidth = 0.1
  ) +
  scale_fill_manual(values = c("Pelagic" = "#82e1da", 
                               "Benthic" = "#103f4d")) +
  labs(fill = "Position") +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(hjust = 0.5, size = 7),
        # legend.direction = "horizontal",
        legend.position = "inside",
        legend.title.position = "top",
        legend.position.inside = c(0.1, 0.8),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"))

# Extract legend
posLegend <- get_legend(phyloTreePosition)

phyloTreePosition
```

Tracking data tree.

```{r}
phyloTreeTag <- phyloTreeOrder +
  geom_fruit(
    data = sharkTraits,
    geom = geom_tile,
    mapping = aes(y = Genus_species, x = 0, 
                  fill = Tracked),
    stat = "identity",
    width = 0.01,
    pwidth = 0.1
  ) +
  scale_fill_manual(values = c("Yes" = "black",
                               "No" = "white")) +
  labs(fill = "Tagged") +
  theme(legend.text = element_text(size = 6),
        legend.title = element_text(hjust = 0.5, size = 7),
        # legend.direction = "horizontal",
        legend.position = "inside",
        legend.title.position = "top",
        legend.position.inside = c(0.1, 0.8),
        legend.key.size = unit(0.3, "cm"),
        legend.background = element_rect(fill = "transparent"))

# Extract legend
tagLegend <- get_legend(phyloTreeTag)

phyloTreeTag
```

Create the mega tree.

```{r}
# Build tree without legends
phyloTreeFull <- phyloTreeOrder +
  geom_fruit(
    data = sharkTraits,
    geom = geom_tile,
    mapping = aes(y = Genus_species, x = 0,
                  fill = Tracked),
    stat = "identity",
    width = 0.01,
    pwidth = 0.07
  ) +
  scale_fill_manual(values = c("Yes" = "black",
                               "No" = "white")) +
  new_scale_fill() +
    geom_fruit(
    data = sharkTraits,
    geom = geom_tile,
    mapping = aes(y = Genus_species, x = 0,
                  fill = WaterColumnPosition),
    stat = "identity",
    width = 0.01,
    pwidth = 0.07,
    offset = 0.1
  ) +
  scale_fill_manual(values = c("Pelagic" = "#82e1da",
                               "Benthic" = "#103f4d")) +
  new_scale_fill() +
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = log(MaximumBodySize), fill = log(MaximumBodySize)),
    orientation = "y",
    stat = "identity",
    pwidth = 0.15,
    offset = 0.15
  ) +
  scale_fill_scico(palette = "bam") +
  new_scale_fill() +
  geom_fruit(
    data = sharkTraits,
    geom = geom_bar,
    mapping = aes(y = Genus_species, x = CFAR, fill = CFAR),
    orientation = "y",
    stat = "identity",
    pwidth = 0.15,
    offset = 0.1
  ) +
  scale_fill_scico(palette = "cork") +
  labs(fill = "") +
  annotate("text", x = 0.54, y = 447, label = "Carcharhiniformes", size = 2.5) +
  annotate("text", x = 0.51, y = 6, label = "Echinorhiniformes", size = 2.5) +
  annotate("text", x = 0.76, y = 133, label = "Heterodontiformes", size = 2.5) +
  annotate("text", x = 0.50, y = 2, label = "Hexanchiformes", size = 2.5) +
  annotate("text", x = 0.78, y = 187, label = "Lamniformes", size = 2.5) +
  annotate("text", x = 0.72, y = 175, label = "Orectolobiformes", size = 2.5) +
  annotate("text", x = 0.51, y = 15, label = "Pristiophoriformes", size = 2.5) +
  annotate("text", x = 0.82, y = 124, label = "Squaliformes", size = 2.5) +
  theme(legend.position = "none",
        margins = unit(c(1.1, 0.5, 0.1, 0),"cm"))

phyloTreeFull

# Add legends in custom position
phyloTreeFinal <- ggdraw() +
  draw_plot(phyloTreeFull) +
  draw_plot(tagLegend, x = 0.51, y = 0.175) +
  draw_plot(posLegend, x = 0.575, y = 0.175) +
  draw_plot(sizeLegend, x = 0.668, y = -0.027) +
  draw_plot(CFARlegend, x = 0.80, y = -0.027)

phyloTreeFinal
```

```{r}
ggsave("Results/Figures/Fig2.tiff", phyloTreeFinal, units = "px",
       device = "tiff",
       width = 2000, height = 2350, dpi = 300)
```

