---
title: "4_brmsAnalysis"
output: html_document
---

## Purpose

To model adjustment sets using the 'brms' package.

Adjustment sets come from dagitty.

### Step 1: Load libraries

Load the libraries used.

```{r}
# Load tidyverse for sanity
library(tidyverse)

# Load ape, geiger, and phytools for dealing with phylogenetic trees
library(ape)
library(geiger)
library(phytools)

# Load brms for analysis
library(brms)

# Load tidybayes for analyzing brms
library(tidybayes)

# Load DHARMa for visualizing model fit
library(DHARMa)

# Load ggtext and gridExtra  for plotting
library(ggtext)
library(cowplot)
library(gridExtra)
library(grid)
```

## Step 2: Load data and phylogenetic trees

```{r}
## Trait Data ##
sharkData <- read.csv("Data/SharkData/TraitData.csv") %>%
  # See 1_Supplemental_CategoricalAnalysis for Benthopelagic > Benthic
  mutate(WaterColumnPosition = case_when(
   str_detect(WaterColumnPosition, "Pelagic") ~ "Pelagic",
   WaterColumnPosition == "Benthic" ~ "Benthic",
   WaterColumnPosition == "Benthopelagic" ~ "Benthic")) %>%
  mutate(WaterColumnPosition = as.factor(WaterColumnPosition))

# Store the original data frame for use later
OGsharkData <- sharkData

## Phylogenetic Tree ##
# Read in all 500 trees
allTrees <- read.nexus("Data/SharkData/PhyloTrees.nex")
```

## Step 3: Prepare data and tree

```{r}
# Set the seed for reproducibility
set.seed(125)

# Sample out a single tree
singleTree <- sample(allTrees, 1)[[1]]

# Check if/where the tree and trait data differ (should return "OK")
nameDrops <- name.check(singleTree, sharkData, sharkData$Genus_species)

# Check length of df = length of tree (should return TRUE)
length(sharkData$Genus_species) == length(singleTree$tip.label)

# Reorder data to match tree just in case
target <- singleTree$tip.label

# Reorder sharkData rows in the same order as singleTree$tip.label
sharkData <- sharkData[match(target, sharkData$Genus_species), ]

# Rename rows to match tip labels
rownames(sharkData) <- sharkData$Genus_species

# Log MaximumBodySize
sharkData <- sharkData %>% 
  mutate(MaximumBodySize = log(MaximumBodySize))

# Set variables to be scaled
vars_to_scale <- c("MaximumBodySize", "ProportionalGillAverage", "AvgTemperature")

# Center and scale continuous variables (vars_to_scale)
sharkData <- sharkData %>%
  mutate(across(all_of(vars_to_scale), scale))
```

## Step 4: Run brms

```{r}
# Create variance covariance matrix
B <- ape::vcv(singleTree)

## Adjustment set 1: AvgTemperature as exposure (CFAR ~ AvgTemperature)
# Build and run model
avgT <- brm(CFAR ~ AvgTemperature + (1|gr(Genus_species, cov = B)),
          data = sharkData,
          data2 = list(B = B),
          family = Gamma(link = "log"),
          warmup = 1000, iter = 3000,
          cores = 2,
          save_pars = save_pars(all = TRUE))

summary(avgT)

# Visualize model fit with DHARMa
model.check.avgT <- createDHARMa(
  simulatedResponse = t(posterior_predict(avgT)),
  observedResponse = sharkData$CFAR,
  fittedPredictedResponse = apply(t(posterior_epred(avgT)), 1, mean),
  integerResponse = FALSE
)

plot(model.check.avgT)

# Save brms object
saveRDS(avgT, "Results/brmsOutputs/brmsAvgT.rds")

## Adjustment set 2: MaximumBodySize as exposure (CFAR ~ MaximumBodySize + WaterColumnPosition)
# Build and run model
maxBodySize <- brm(CFAR ~ MaximumBodySize + WaterColumnPosition +
                          (1|gr(Genus_species, cov = B)),
          data = sharkData,
          data2 = list(B = B),
          family = Gamma(link = "log"),
          warmup = 1000, iter = 3000,
          cores = 2,
          save_pars = save_pars(all = TRUE))

summary(maxBodySize)

# Visualize model fit with DHARMa
model.check.maxBodySize <- createDHARMa(
  simulatedResponse = t(posterior_predict(maxBodySize)),
  observedResponse = sharkData$CFAR,
  fittedPredictedResponse = apply(t(posterior_epred(maxBodySize)), 1, mean),
  integerResponse = FALSE
)

plot(model.check.maxBodySize)

# Save brms object
saveRDS(maxBodySize, "Results/brmsOutputs/brmsMaxBodySize.rds")

## Adjustment set 3: WaterColumnPosition as exposure 
# (CFAR ~ WaterColumnPosition)
# Build and run model
waterColPos <- brm(CFAR ~ 0 + WaterColumnPosition +
                     (1|gr(Genus_species, cov = B)),
          data = sharkData,
          data2 = list(B = B),
          family = Gamma(link = "log"),
          warmup = 1000, iter = 3000,
          cores = 2,
          save_pars = save_pars(all = TRUE))

summary(waterColPos)

# Visualize model fit with DHARMa
model.check.waterColPos <- createDHARMa(
  simulatedResponse = t(posterior_predict(waterColPos)),
  observedResponse = sharkData$CFAR,
  fittedPredictedResponse = apply(t(posterior_epred(waterColPos)), 1, mean),
  integerResponse = FALSE
)

plot(model.check.waterColPos)

# Save brms object
saveRDS(waterColPos, "Results/brmsOutputs/brmsWaterColPos.rds")

## Adjustment set 4: ProportionalGillAverage as exposure 
# (CFAR ~ ProportionalGillAverage + MaximumBodySize + WaterColumnPosition)
# Build and run model
gillHeight <- brm(CFAR ~ ProportionalGillAverage + MaximumBodySize +
                    WaterColumnPosition +
                    (1|gr(Genus_species, cov = B)),
          data = sharkData,
          data2 = list(B = B),
          family = Gamma(link = "log"),
          warmup = 1000, iter = 3000,
          cores = 2,
          save_pars = save_pars(all = TRUE))

summary(gillHeight)

# Visualize model fit with DHARMa
model.check.gillHeight <- createDHARMa(
  simulatedResponse = t(posterior_predict(gillHeight)),
  observedResponse = sharkData$CFAR,
  fittedPredictedResponse = apply(t(posterior_epred(gillHeight)), 1, mean),
  integerResponse = FALSE
)

plot(model.check.gillHeight)

# Save brms object
saveRDS(gillHeight, "Results/brmsOutputs/brmsGillHeight.rds")
```

## Optional: Load brms objects if revisiting script

```{r}
# Load brms objects using readRDS
avgT <- readRDS("Results/brmsOutputs/brmsAvgT.rds")
maxBodySize <- readRDS("Results/brmsOutputs/brmsMaxBodySize.rds")
waterColPos <- readRDS("Results/brmsOutputs/brmsWaterColPos.rds")
gillHeight <- readRDS("Results/brmsOutputs/brmsGillHeight.rds")
```

## Step 5: Analyze brms results

```{r}
# Continuous variables
continuous_list <- list(
  # Average temperature model
  avgT %>% spread_draws(b_AvgTemperature) %>%
    transmute(estimate = b_AvgTemperature, term = "Average Temperature", 
              model = "avgT"),
  # Body Size model
  maxBodySize %>% spread_draws(b_MaximumBodySize) %>%
    transmute(estimate = b_MaximumBodySize, term = "Maximum Body Size", 
              model = "maxBodySize"),
  # Gill slit height model
  gillHeight %>% spread_draws(b_ProportionalGillAverage) %>%
    transmute(estimate = b_ProportionalGillAverage, term = "Gill Slit Height", 
              model = "gillHeight")
)

# Water column position 
waterPos_list <- list(
  waterColPos %>% spread_draws(b_WaterColumnPositionBenthic) %>%
    transmute(estimate = exp(b_WaterColumnPositionBenthic), term = "Benthic", 
              model = "waterColPos"),
  waterColPos %>% spread_draws(b_WaterColumnPositionPelagic) %>%
    transmute(estimate = exp(b_WaterColumnPositionPelagic), term = "Pelagic", 
              model = "waterColPos") 
)

# Function to summarize each coefficient
summarize_coefs <- function(df) {
  df %>%
    # Mean + 95% credible interval
    mean_qi(estimate, .width = 0.95)
}

# Bind rows by column names to combine all spread_draws
continuousCoefs <- bind_rows(continuous_list)
waterCoefs <- bind_rows(waterPos_list)

# Run continuousCoefs and waterCoefs through summarize_coefs function
continuousSummary <- continuousCoefs %>% 
  summarize_coefs()

waterPosSummary <- waterCoefs %>%
  summarize_coefs()
```

## Step 6: Plot brms results

```{r}
## Continuous variables ##
# Create a tibble of labels
contLabels <- tibble::tibble(
  term = c("Maximum Body Size", "Gill Slit Height", "Average Temperature"),
  label = c(
    # Maximum Body Size
    "<b><span style='color:#424B54;'>Maximum Body Size</span></b> + Water Column Position",
    
    # Gill Slit Height
    "<b><span style='color:#93A8AC;'>Gill Slit Height</span></b> + Maximum Body Size + Water Column Position",
    
    # Avg Temp
    "<b><span style='color:#9B6A6C;'>Average Temperature</span></b>"
  )
)

# Compute centers
contCenters <- continuousCoefs %>%
  group_by(term) %>%
  summarize(estimate = mean(estimate), .groups = "drop") %>%
  mutate(y = fct_rev(fct_reorder(term, estimate, .desc = TRUE)))

# Join centers to labels
contLabels <- contLabels %>% 
  left_join(contCenters, by = "term")

# Plot the continuous variables
contPlot <- ggplot(continuousCoefs, 
                   aes(x = estimate, 
                       y = fct_rev(fct_reorder(term, estimate, .desc = TRUE)),
                       fill = term)) +
  stat_halfeye(.width = c(0.95, 0.95, 0.95), point_interval = mean_qi,
               scale = .82) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = c("Gill Slit Height" = "#93A8AC",
                               "Average Temperature" = "#9B6A6C",
                               "Maximum Body Size" = "#424B54")) +
  xlim(-0.45, 0.7) +
  scale_y_discrete(expand = expansion(mult = c(0.15, 0))) +
  labs(x = "Coefficient Estimate",
       title = "A") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        plot.margin = unit(c(0, 0.25, 0, 0.15), "cm")) +
  geom_richtext(
    data = contLabels,
    aes(x = estimate, y = term, label = label),
    inherit.aes = FALSE,
    hjust = 0.5,
    size = 3.4,
    vjust = 1.3,
    fill = "white",
    label.color = NA
  ) +
  geom_segment(aes(x = -0.1093399, xend = -0.1093399, y = 1.03, yend = 1.7), 
               color = "#FFFFFF") +
  geom_segment(aes(x = 0.1366995, xend = 0.1366995, y = 2.03, yend = 2.82), 
               color = "#FFFFFF") +
  geom_segment(aes(x = 0.1527054, xend = 0.1527054, y = 3.03, yend = 3.8), 
               color = "#FFFFFF")

contPlot
```

```{r}
## Water column position ##
# Create a tibble of labels
waterPosLabels <- tibble::tibble(
  term = factor(c("Benthic", "Pelagic"),
                levels = c("Benthic", "Pelagic")),
  label = c(
    "<b><span style='color:#103f4d;'>Benthic</span></b>",
    "<b><span style='color:#82e1da;'>Pelagic</span></b>"
  )
)

# Compute centers
waterPosCenters <- waterCoefs %>%
  group_by(term) %>%
  summarize(estimate = mean(estimate), .groups = "drop") %>%
  mutate(y = fct_rev(fct_reorder(term, estimate, .desc = TRUE)))

# Join centers to labels
waterPosLabels <- waterPosLabels %>% 
  left_join(waterPosCenters, by = "term")

waterPosTitle <- tibble::tibble(
  term = "Title",
  label = c(
    "<b>Water Column Position</b> + Average Temperature + Maximum Body Size"))

# Plot water column position
waterPosPlot <- ggplot(waterCoefs, 
                   aes(x = estimate, 
                       y = fct_rev(fct_reorder(term, estimate, .desc = TRUE)),
                       fill = term)) +
  stat_halfeye(.width = c(0.95, 0.95, 0.95), point_interval = mean_qi) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  scale_fill_manual(values = c("Benthic" = "#103f4d",
                               "Pelagic" = "#82e1da")) +
  scale_x_continuous(limits = c(-0.1, 5.5),
                     breaks = seq(0, 5, 1)) +
  labs(title = "B",
       subtitle = "**Water Column Position**",
       x = "Estimated Marginal Means") +
  theme_classic() +
  theme(plot.subtitle = element_markdown(margin = margin(t = 10, b = -17), 
                                      hjust = 0.5, size = 9),
        axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        plot.margin = unit(c(-0.1, 0.25, 0, 0.15), "cm")) +
  geom_richtext(
    data = waterPosLabels,
    aes(x = estimate, y = term, label = label),
    inherit.aes = FALSE,
    hjust = 0.5,
    size = 3.4,
    vjust = 1.3,
    fill = NA,
    label.color = NA
  )

waterPosPlot
```

* Warning about removed rows: 
The very long, improbable tails of water col position are trimmed by the x-axis limits

Arrange brms results plots.

```{r}
spacer <- nullGrob()

brms_left_column <- arrangeGrob(spacer, contPlot, spacer, waterPosPlot, spacer,
                                ncol = 1, heights = c(0.05, 2.15, 0.1, 0.85, 0.05))
```

* Warning about removed rows: 
The very long, improbable tails of water col position are trimmed by the x-axis limits

# Step 7: Scatter plots of variables

Plot data vs. CFAR and add brms 

```{r}
## Maximum Body Size ##
# Tibble of results
maxSize <- tibble(
   MaximumBodySize = seq(
    min(sharkData$MaximumBodySize, na.rm = TRUE),
    max(sharkData$MaximumBodySize, na.rm = TRUE),
    # Smooth the line
    length.out = 200)
  )

# Create prediction grid with both water column positions
maxSize_avg <- maxSize %>%
  crossing(
    WaterColumnPosition = levels(sharkData$WaterColumnPosition)
  )

# Expected values
maxBodySize_epred_draws <- add_epred_draws(newdata = maxSize_avg, 
                                           object = maxBodySize,
                                           ndraws = 200, re_formula = NA)

# Average to get one y per x per draw
maxBody_epred_avg <- maxBodySize_epred_draws %>%
  group_by(.draw, MaximumBodySize) %>%
  summarise(
    .epred = mean(.epred),
    .groups = "drop"
  )

# Plot
maxBodySize_plot <- ggplot(maxBody_epred_avg, aes(x = MaximumBodySize, y = .epred)) +
  stat_lineribbon(.width = 0, color = "#20272f", alpha = 0.5) +  
  geom_line(aes(group = .draw), color = "#20272f", linewidth = 0.1, alpha = 0.1) +
  geom_point(data = sharkData, aes(x = MaximumBodySize, y = CFAR),
             alpha = 0.7, color = "#424B54", size = 2) +
  labs(title = "C",
       x = "Maximum Body Size", y = "CFAR") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        plot.margin = unit(c(0, 0.1, 0, 0), "cm"))

maxBodySize_plot

# Because scaling on the x axis
log_body <- log(OGsharkData$MaximumBodySize)

body_mean <- mean(log_body, na.rm = TRUE)
body_sd   <- sd(log_body, na.rm = TRUE)

# Modifying the breaks
BodyScaled_breaks <- c(-2, 0, 2, 4)

BodyLog_breaks <- BodyScaled_breaks * body_sd + body_mean

BodyOriginal_breaks <- exp(BodyLog_breaks)
BodyOriginal_breaks
  
maxBodySize_plot <- ggplot(maxBody_epred_avg, aes(x = MaximumBodySize, y = .epred)) +
  geom_line(aes(group = .draw), color = "#20272f", linewidth = 0.1, alpha = 0.1) +
  geom_point(data = sharkData, aes(x = MaximumBodySize, y = CFAR),
             alpha = 0.7, color = "#424B54", size = 2) +
  stat_lineribbon(.width = 0, color = "#20272f", alpha = 0.5) +
  scale_x_continuous(breaks = BodyScaled_breaks,
                     labels = round(BodyOriginal_breaks)) +
  labs(title = "C",
       x = "Maximum Body Size (cm)", y = "CFAR") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        plot.margin = unit(c(0, 0.1, 0, 0), "cm"))

maxBodySize_plot

## Gill Slit Height ##
# Tibble of results
gillSlitHeight <- tibble(
  ProportionalGillAverage = seq(
    min(sharkData$ProportionalGillAverage, na.rm = TRUE),
    max(sharkData$ProportionalGillAverage, na.rm = TRUE),
    length.out = 200))

# Create a mean Max Body Size to "hold"
mean_maxSize <- mean(sharkData$MaximumBodySize)

# Create prediciton grid with both water column positions and max body size
gillSlit_avg <- gillSlitHeight %>%
  crossing(WaterColumnPosition = unique(sharkData$WaterColumnPosition)) %>%
  mutate(
    MaximumBodySize = mean_maxSize,
    WaterColumnPosition = factor(
      WaterColumnPosition,
      levels = levels(as.factor(sharkData$WaterColumnPosition))
    )
  )

# Expected values
gillSlit_epred_draws <- add_epred_draws(newdata = gillSlit_avg, object = gillHeight,
                                           ndraws = 200, re_formula = NA)

# Average to get one y per x per draw
gillSlit_epred_avg <- gillSlit_epred_draws %>%
  group_by(.draw, ProportionalGillAverage) %>%
  summarise(
    .epred = mean(.epred),
    .groups = "drop"
  )

# Plot
gillHeight_plot <- ggplot(gillSlit_epred_avg, aes(x = ProportionalGillAverage, y = .epred)) +
  geom_line(aes(group = .draw), color = "#56696d", linewidth = 0.1, alpha = 0.1) +
  geom_point(data = sharkData, aes(x = ProportionalGillAverage, y = CFAR),
             alpha = 0.7, color = "#93A8AC", size = 2) +
  stat_lineribbon(.width = 0, color = "#56696d", alpha = 0.5) +
  labs(title = "D",
       x = "Gill Slit Height (% TL)", y = "CFAR") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        plot.margin = unit(c(0, 0.1, 0, 0), "cm"))
  
gillHeight_plot

# Because scaling on the x axis
gill_mean <- mean(OGsharkData$ProportionalGillAverage, na.rm = TRUE)
gill_sd   <- sd(OGsharkData$ProportionalGillAverage, na.rm = TRUE)

# Modifying the breaks
GillScaled_breaks <- c(-2, 0, 2, 4)

GillLog_breaks <- GillScaled_breaks * gill_sd + gill_mean

GillOriginal_breaks <- exp(GillLog_breaks)
GillOriginal_breaks

# Plot
gillHeight_plot <- ggplot(gillSlit_epred_avg, aes(x = ProportionalGillAverage, y = .epred)) +
  geom_line(aes(group = .draw), color = "#56696d", linewidth = 0.1, alpha = 0.1) +
  geom_point(data = sharkData, aes(x = ProportionalGillAverage, y = CFAR),
             alpha = 0.8, color = "#93A8AC", size = 2) +
  stat_lineribbon(.width = 0, color = "#56696d", alpha = 0.5) + 
  scale_x_continuous(
    breaks = GillScaled_breaks,
    labels = round(GillOriginal_breaks, 2)
  ) +
  labs(title = "D",
       x = "Gill Slit Height (% TL)", y = "CFAR") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        plot.margin = unit(c(0, 0.1, 0, 0), "cm"))
  
gillHeight_plot

## Average Temperature ##
# Tibble of results
avgTemp <- tibble(
  AvgTemperature = seq(
    min(sharkData$AvgTemperature, na.rm = TRUE),
    max(sharkData$AvgTemperature, na.rm = TRUE),
    # Smooth the line
    length.out = 200)
  )

# Expected values
avgT_epred_draws <- add_epred_draws(avgTemp, avgT, ndraws = 100, 
                                    re_formula = NA)

# Plot
avgT_plot <- ggplot(avgT_epred_draws, aes(x = AvgTemperature, y = .epred)) +
  geom_line(aes(group = .draw), color = "#5d3233", linewidth = 0.1, alpha = 0.1) +
  geom_point(data = sharkData, aes(x = AvgTemperature, y = CFAR),
             alpha = 0.8, color = "#9B6A6C", size = 2) +
  stat_lineribbon(.width = 0, color = "#5d3233", alpha = 0.5) + 
  labs(title = "E",
       x = "Average Temperature (\u00B0C)", y = "CFAR") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        plot.margin = unit(c(0, 0.1, 0, 0), "cm"))
  
avgT_plot

# Because scaling on the x axis
temp_mean <- mean(OGsharkData$AvgTemperature, na.rm = TRUE)
temp_sd   <- sd(OGsharkData$AvgTemperature, na.rm = TRUE)

# Modifying the breaks
TempScaled_breaks <- c(-2, -1, 0, 1, 2)

TempLog_breaks <- TempScaled_breaks * temp_sd + temp_mean

TempOriginal_breaks <- TempLog_breaks
TempOriginal_breaks

# Plot
avgT_plot <- ggplot(avgT_epred_draws, aes(x = AvgTemperature, y = .epred)) +  
  geom_line(aes(group = .draw), color = "#5d3233", linewidth = 0.1, alpha = 0.1) +
  geom_point(data = sharkData, aes(x = AvgTemperature, y = CFAR),
             alpha = 0.7, color = "#9B6A6C", size = 2) +
  stat_lineribbon(.width = 0, color = "#5d3233", alpha = 0.5) +
  scale_x_continuous(
    breaks = TempScaled_breaks,
    labels = round(TempOriginal_breaks, 1)
  ) +
  labs(title = "E",
       x = "Average Temperature (\u00B0C)", y = "CFAR") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        plot.margin = unit(c(0, 0.1, 0, 0), "cm"))
  
avgT_plot

## Water Column Position ##
waterCol_plot <- ggplot(data = sharkData, aes(x = WaterColumnPosition, y = CFAR, 
                                              fill = WaterColumnPosition, 
                                              color = WaterColumnPosition)) + 
  geom_violin(alpha = 0.9) +
  scale_fill_manual(values = c("Benthic" = "#103f4d",
                               "Pelagic" = "#82e1da")) +
  scale_color_manual(values = c("Benthic" = "#103f4d",
                               "Pelagic" = "#82e1da")) +
  labs(title = "F",
       x = element_blank(), y = "CFAR") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 8),
        axis.title.x = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.y = element_text(size = 10),
        plot.margin = unit(c(0, 0.1, 0, 0), "cm"))

waterCol_plot
```

Arrange scatter plots and violin plot.

```{r}
spacer <- nullGrob()

scatter_right_column <- arrangeGrob(spacer, maxBodySize_plot, spacer,
                                    gillHeight_plot, spacer, avgT_plot, 
                                    spacer, waterCol_plot,
                                    ncol = 1, 
                                    heights = c(0.05, 1, 0.01, 1, 0.01,
                                                          1, 0.06, 1.21, 0.125))
```

```{r}
# Combine plots
Figure4 <- arrangeGrob(
  brms_left_column,
  spacer,
  scatter_right_column,
  ncol = 3,
  widths = c(1.2, 0.05, 0.8)
)

# Save combined plot
ggsave("Results/Figures/Fig4.tiff", 
       plot = Figure4, units = "px",
       device = "tiff",
       width = 2250, height = 2500)
```
