---
title: "4_mappingCFAR"
output: html_document
---

## Purpose

To map CFAR per hex cell.

## Step 1: Load libraries

Load the libraries used.

```{r}
# Load tidyverse
library(tidyverse)

# Load chondrichthyes.species.ranges
# Private package (https://github.com/JayMatsushiba)
library(chondrichthyes.species.ranges)

# Load littany of packages for mapping
library(sf)
library(patchwork)
library(rnaturalearthdata)
library(rnaturalearth)
library(furrr)
library(purrr)
library(tictoc)

# Load scico for color palette
library(scico)

# Load libraries for combining plots
library(cowplot)
library(gridExtra)
library(grid)
```

## Step 2: Load data, hex cells, and species ranges

```{r}
# Load shark data
sharkData <- read.csv("Data/SharkData/TraitData.csv") %>%
  # Spatial data is formatted as "Genus species" instead of "Genus_species"
  mutate(Genus_species = str_replace_all(Genus_species, "_", " "))

# Load hex cells
hex <- st_read("Data/HexCells")

# Load species ranges
ranges <- chondrichthyes.species.ranges::get_ranges() %>%
  # Reproject ranges onto the same CRS as the hex grid
  st_transform(st_crs(hex))
```

## Step 3: Calculate intersection between ranges and hex grid

```{r}
# Calculate intersection between ranges and hex grid
# tic() starts the timer
tic("This is how long this process takes")
# st_intersects() calculates the intersection between species ranges and the hex grid
range_intersect_hex <- st_intersects(ranges, hex)
# toc() ends the timer
toc()
```

## Step 4: Join CFAR data onto range data

```{r}
# Keep only the species names and one trait
CFAR_data <- sharkData %>%
  select(sci_name = Genus_species, CFAR)

# Left join CFAR data to range data
allRanges <- left_join(ranges, CFAR_data, by = "sci_name")
```

## Step 5: Calculate CFAR per hex cell (mean and variability)

```{r}
# Create a long table of species-hex links from list (output of st_intersects())
range_hex_df <- tibble(
  # rep() to repeat each species index per intersecting hex
  # lengths() determines how many hexes each species overlaps
  species_index = rep(seq_along(range_intersect_hex), lengths(range_intersect_hex)),
  # Flattens hex indices into one vector
  hex_index = unlist(range_intersect_hex)
)

# Use species_index to index ranges$CFAR
range_hex_df <- range_hex_df %>%
  # Add a CFAR value to every species-hex combo
  mutate(CFAR = allRanges$CFAR[species_index])

# Collapse species-level data into hex-level summaries
hex_cfar_stats <- range_hex_df %>%
  # Group_by() hex (all species that occur in the same hex)
  group_by(hex_index) %>%
  # Compute summary statistics: mean and var
  summarise(
    mean_CFAR = mean(CFAR, na.rm = TRUE),
    var_CFAR = var(CFAR, na.rm = TRUE),
    # Ungroup results
    .groups = "drop"
  )

# Add CFAR mean and var to hex
hex$mean_CFAR <- NA_real_
hex$var_CFAR <- NA_real_

# Place values into the correct rows
hex$mean_CFAR[hex_cfar_stats$hex_index] <- hex_cfar_stats$mean_CFAR
hex$var_CFAR[hex_cfar_stats$hex_index] <- hex_cfar_stats$var_CFAR
```

## Step 6: Map Mean CFAR and Variability in CFAR

Create a world for plotting.

```{r}
# Create world
world <- ne_countries(scale = "medium", returnclass = "sf")
# Match hex's CRS (Cylindrical Equal Area)
world <- st_transform(world, st_crs(hex))

# Define bounding box from world 
world_extent <- sf::st_bbox(world)
world_extent$xmin # -20037508 
world_extent$xmax # 20037508
world_extent$ymin # -6363885 
world_extent$ymax # 6323867
```

Plot CFAR.

```{r}
mean_CFAR_map <- ggplot() +
  geom_sf(data = world, fill = "grey82", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = mean_CFAR), color = NA) +
  scale_fill_scico(palette = "corkO", na.value = "grey95", name = "Mean CFAR",
                   # Limits consistent across plots
                   limits = c(0.15, 5.53)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 7),
    plot.subtitle = element_text(face = "bold", size = 8),
    legend.position = "none",
    plot.margin = unit(c(-0.1, 0, 0, 0), "cm")
  ) +
  labs(subtitle = "a",
       y = "All Sharks")

mean_CFAR_map

var_CFAR_map <- ggplot() +
  geom_sf(data = world, fill = "grey80", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = var_CFAR), color = NA) +
  scale_fill_gradient(low = "#e0aaff", high = "#311e4f", na.value = "grey95", 
                      name = "Variation in CFAR",
                      # Limits consistent across plots
                      limits = c(0, 5)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) + labs(subtitle = "b") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.subtitle = element_text(face = "bold", size = 8),
    legend.position = "none",
    plot.margin = unit(c(-0.1, 0, 0, 0), "cm")
  )

var_CFAR_map
```

Extract legends.

```{r}
# Extract legend
extract_legend <- function(p) {
  g <- ggplotGrob(p)
  legend <- g$grobs[
    sapply(g$grobs, function(x) x$name) == "guide-box"
  ][[1]]
  return(legend)
}

# Left side legend
legend_plot_left <- ggplot() +
  geom_sf(data = world, fill = "grey82", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = mean_CFAR), color = NA) +
  scale_fill_scico(palette = "corkO", na.value = "grey95", 
                   name = "Mean CFAR",
                   # Limits consistent across plots
                   limits = c(0.15, 5.53)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) +
  labs(title = "a",
       y = "All Sharks") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title.x = element_blank(),,
    axis.title.y = element_text(size = 7),
    plot.title = element_text(face = "bold", size = 8),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "transparent"),
    legend.key.width = unit(1.8, 'cm'),
    legend.title = element_text(size = 7, hjust = 0.5),
    legend.title.position = "top",
    legend.text = element_text(size = 6),
    plot.margin = unit(c(0, 0, 0, 0), "cm")
  )

legend_left <- extract_legend(legend_plot_left)

# Right side legend
legend_plot_right <- ggplot() +
  geom_sf(data = world, fill = "grey80", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = var_CFAR), color = NA) +
  scale_fill_gradient(low = "#e0aaff", high = "#311e4f", na.value = "grey95", 
                      name = "Variation in CFAR",
                      # Limits consistent across plots
                      limits = c(0, 5)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)) +
  labs(subtitle = "b") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.subtitle = element_text(face = "bold", size = 8),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "transparent"),
    legend.key.width = unit(1.8, 'cm'),
    legend.title = element_text(size = 7, hjust = 0.5),
    legend.title.position = "top",
    legend.text = element_text(size = 6),
    plot.margin = unit(c(0, 0, 0, 0), "cm")
  )

legend_right <- extract_legend(legend_plot_right)
```

# Plot CFAR for each habitat use (PrimaryHabitat) category: pelagic, coastal, deepwater.

1. Pelagic species

```{r}
# Filter sharkData for species by habitat type
pelagic_sharks <- sharkData %>%
  filter(PrimaryHabitat == "Pelagic")

# Select species name and CFAR
pelagicCFAR <- pelagic_sharks %>%
  select(sci_name = Genus_species, CFAR)

# Left join pelagic CFAR data to range data
ranges_pelagic <- ranges %>%
  left_join(pelagicCFAR, by = "sci_name") %>%
  filter(!is.na(CFAR))

# Timer for how long it takes to intersects ranges
tic("Pelagic species: range–hex intersection")
range_intersect_hex_pelagic <- st_intersects(ranges_pelagic, hex)
toc()

# Create tibble
range_hex_df_pelagic <- tibble(
  species_index = rep(
    seq_along(range_intersect_hex_pelagic),
    lengths(range_intersect_hex_pelagic)
  ),
  hex_index = unlist(range_intersect_hex_pelagic)
)

range_hex_df_pelagic <- range_hex_df_pelagic %>%
  mutate(CFAR = ranges_pelagic$CFAR[species_index])

hex_cfar_stats_pelagic <- range_hex_df_pelagic %>%
  group_by(hex_index) %>%
  summarise(
    mean_CFAR_pelagic = mean(CFAR, na.rm = TRUE),
    var_CFAR_pelagic  = var(CFAR,  na.rm = TRUE),
    .groups = "drop"
  )

hex$mean_CFAR_pelagic <- NA_real_
hex$var_CFAR_pelagic  <- NA_real_

hex$mean_CFAR_pelagic[hex_cfar_stats_pelagic$hex_index] <-
  hex_cfar_stats_pelagic$mean_CFAR_pelagic

hex$var_CFAR_pelagic[hex_cfar_stats_pelagic$hex_index] <-
  hex_cfar_stats_pelagic$var_CFAR_pelagic

# Plot pelagic species
mean_pelagic_map <- ggplot() +
  geom_sf(data = world, fill = "grey80", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = mean_CFAR_pelagic), color = NA) +
  
  scale_fill_scico(palette = "corkO", na.value = "grey95", name = "CFAR",
                   # Limits consistent across plots
                   limits = c(0.15, 5.53)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 7),
    plot.title.position = "panel",
    plot.title = element_text(face = "bold", size = 8),
    legend.position = "none",
    plot.margin = unit(c(0, 0, 0, 0), "cm")
  ) +
  labs(title = "c",
       y = "Pelagic Sharks")

mean_pelagic_map

var_pelagic_map <- ggplot() +
  geom_sf(data = world, fill = "grey80", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = var_CFAR_pelagic), color = NA) +
  scale_fill_gradient(low = "#e0aaff", high = "#311e4f", na.value = "grey95", name = "Var",
                      # Limits consistent across plots
                      limits = c(0, 5)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.title.position = "panel",
    plot.title = element_text(face = "bold", size = 8),
    legend.position = "none",
    plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  labs(title = "d")

var_pelagic_map
```

2. Coastal species

```{r}
coastal_sharks <- sharkData %>%
  filter(PrimaryHabitat == "Coastal")

# Select species name and CFAR
coastalCFAR <- coastal_sharks %>%
  select(sci_name = Genus_species, CFAR)

# Left join coastal CFAR data to range data
ranges_coastal <- ranges %>%
  left_join(coastalCFAR, by = "sci_name") %>%
  filter(!is.na(CFAR))

# Timer for how long it takes to intersects ranges
tic("Coastal species: range–hex intersection")
range_intersect_hex_coastal <- st_intersects(ranges_coastal, hex)
toc()

range_hex_df_coastal <- tibble(
  species_index = rep(
    seq_along(range_intersect_hex_coastal),
    lengths(range_intersect_hex_coastal)
  ),
  hex_index = unlist(range_intersect_hex_coastal)
)

range_hex_df_coastal <- range_hex_df_coastal %>%
  mutate(CFAR = ranges_coastal$CFAR[species_index])

hex_cfar_stats_coastal <- range_hex_df_coastal %>%
  group_by(hex_index) %>%
  summarise(
    mean_CFAR_coastal = mean(CFAR, na.rm = TRUE),
    var_CFAR_coastal  = var(CFAR,  na.rm = TRUE),
    .groups = "drop"
  )

hex$mean_CFAR_coastal <- NA_real_
hex$var_CFAR_coastal  <- NA_real_

hex$mean_CFAR_coastal[hex_cfar_stats_coastal$hex_index] <-
  hex_cfar_stats_coastal$mean_CFAR_coastal

hex$var_CFAR_coastal[hex_cfar_stats_coastal$hex_index] <-
  hex_cfar_stats_coastal$var_CFAR_coastal

# Plot coastal species
mean_coastal_map <- ggplot() +
  geom_sf(data = world, fill = "grey80", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = mean_CFAR_coastal), color = NA) +
  
  scale_fill_scico(palette = "corkO", na.value = "grey95", name = "CFAR",
                   # Limits consistent across plots
                   limits = c(0.15, 5.53)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title.x = element_blank(),,
    axis.title.y = element_text(size = 7),
    plot.title.position = "panel",
    plot.title = element_text(face = "bold", size = 8),
    legend.position = "none",
    plot.margin = unit(c(0, 0, 0, 0), "cm")
  ) +
  labs(title = "e",
       y = "Coastal Sharks")

mean_coastal_map

var_coastal_map <- ggplot() +
  geom_sf(data = world, fill = "grey80", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = var_CFAR_coastal), color = NA) +
  scale_fill_gradient(low = "#e0aaff", high = "#311e4f", na.value = "grey95", name = "Var",
                      # Limits consistent across plots
                      limits = c(0, 5)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.title.position = "panel",
    plot.title = element_text(face = "bold", size = 8),
    legend.position = "none",
    plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  labs(title = "f")

var_coastal_map
```

3. Deepwater species

```{r}
deepwater_sharks <- sharkData %>%
  filter(PrimaryHabitat == "Deepwater")

# Select species name and CFAR
deepwaterCFAR <- deepwater_sharks %>%
  select(sci_name = Genus_species, CFAR)

# Left join deepwater CFAR data to range data
ranges_deepwater <- ranges %>%
  left_join(deepwaterCFAR, by = "sci_name") %>%
  filter(!is.na(CFAR))

# Timer for how long it takes to intersects ranges
tic("Deepwater species: range–hex intersection")
range_intersect_hex_deepwater <- st_intersects(ranges_deepwater, hex)
toc()

range_hex_df_deepwater <- tibble(
  species_index = rep(
    seq_along(range_intersect_hex_deepwater),
    lengths(range_intersect_hex_deepwater)
  ),
  hex_index = unlist(range_intersect_hex_deepwater)
)

range_hex_df_deepwater <- range_hex_df_deepwater %>%
  mutate(CFAR = ranges_deepwater$CFAR[species_index])

hex_cfar_stats_deepwater <- range_hex_df_deepwater %>%
  group_by(hex_index) %>%
  summarise(
    mean_CFAR_deepwater = mean(CFAR, na.rm = TRUE),
    var_CFAR_deepwater  = var(CFAR,  na.rm = TRUE),
    .groups = "drop"
  )

hex$mean_CFAR_deepwater <- NA_real_
hex$var_CFAR_deepwater  <- NA_real_

hex$mean_CFAR_deepwater[hex_cfar_stats_deepwater$hex_index] <-
  hex_cfar_stats_deepwater$mean_CFAR_deepwater

hex$var_CFAR_deepwater[hex_cfar_stats_deepwater$hex_index] <-
  hex_cfar_stats_deepwater$var_CFAR_deepwater

# Plot deepwater species
mean_deepwater_map <- ggplot() +
  geom_sf(data = world, fill = "grey80", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = mean_CFAR_deepwater), color = NA) +
  
  scale_fill_scico(palette = "corkO", na.value = "grey95", name = "CFAR",
                   # Limits consistent across plots
                   limits = c(0.15, 5.53)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title.x = element_blank(),,
    axis.title.y = element_text(size = 7),
    plot.title.position = "panel",
    plot.title = element_text(face = "bold", size = 8),
    legend.position = "none",
    plot.margin = unit(c(0, 0, -0.2, 0), "cm")
  ) +
  labs(title = "g",
       y = "Deepwater Sharks")

mean_deepwater_map

var_deepwater_map <- ggplot() +
  geom_sf(data = world, fill = "grey80", color = "grey90", size = 0.2) +
  geom_sf(data = hex, aes(fill = var_CFAR_deepwater), color = NA) +
  scale_fill_gradient(low = "#e0aaff", high = "#311e4f", na.value = "grey95", name = "Var",
                      # Limits consistent across plots
                      limits = c(0, 5)) +
  # Set x limits and prevent expanding of plot beyond world
  scale_x_continuous(
    limits = c(world_extent$xmin, world_extent$xmax),
    expand = c(0, 0)
  ) +
  # Set y limits and prevent expanding of plot beyond world
  scale_y_continuous(
    limits = c(world_extent$ymin, world_extent$ymax),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.subtitle = element_text(face = "bold", size = 8),
    legend.position = "none",
    plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  labs(subtitle = "h")

var_deepwater_map
```

Combine all plots.

```{r}
spacer <- nullGrob()

left_column <- arrangeGrob(
  legend_left,
  arrangeGrob(mean_CFAR_map, mean_pelagic_map,
              mean_coastal_map, mean_deepwater_map,
              ncol = 1),
  heights = c(0.14, 1)
)

right_column <- arrangeGrob(
  legend_right,
  arrangeGrob(var_CFAR_map, var_pelagic_map,
              var_coastal_map, var_deepwater_map,
              ncol = 1),
  heights = c(0.14, 1)
)

Figure4 <- arrangeGrob(
  spacer,
  left_column,
  spacer,
  right_column,
  ncol = 4,
  widths = c(0.01, 0.9, 0.01, 0.9)
)

ggsave("Results/Figures/Fig4.tiff", 
       plot = Figure4, units = "px",
       device = "tiff",
       bg = "white",
       width = 2250, height = 2000, dpi = 300)
```
